---
title: "Biostat M280 Homework 2 by Huijun An"
subtitle: Due Feb 8 @ 11:59PM
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q1. Travis CI

1. Read the blog post [_Adding CI To RMarkdown Documents_](http://www.goring.org/resources/Adding_CI_To_RMarkdown.html) by Simon Goring. Understand what is continuous integration (CI) and why we need CI.

2. Set up Travis CI for your homework 2 solution, e.g., `hw2sol.Rmd`. Graders will look for a CI passing badge on the `README.Rmd` page of your repository:

  develop branch build status:[![Build Status](https://travis-ci.com/huijunan/biostat-m280-2019-winter.svg?token=zTicFbMBPAEEe9msLo8r&branch=develop)](https://travis-ci.com/huijunan/biostat-m280-2019-winter)
  
  master branch build status: [![Build Status](https://travis-ci.com/huijunan/biostat-m280-2019-winter.svg?token=zTicFbMBPAEEe9msLo8r&branch=master)](https://travis-ci.com/huijunan/biostat-m280-2019-winter)



## Q2. FizzBuzz Test

The FizzBuzz test is a common programming interview question used to establish if a candidate can actually program in a language that they claim experience in. The problem statement is as follows:

> "Write a program that given a list of numbers as input, prints each number on their own line. But for multiples of three print "Fizz" instead of the number and for the multiples of five print “Buzz”. For numbers which are multiples of both three and five print “FizzBuzz”."

1. Your goal here is to implement FizzBuzz as an R function called `fizzbuzz`. Your fizzbuzz function should conform to the description provided above in terms of output and accept an integer argument, which can either be a scalar or vector. As such, your function should correctly print `Fizz`, `Buzz`, `FizzBuzz`, or the actual number for each integer supplied on a separate line. You should program defensively - validate any input and make sure that you have a sane response to any invalid input.

    You must also include a write up of your implementation that broadly describes how you approached the problem and constructed your solution (think something along the lines of the methods section of a journal article).

    This is not a terribly complex or difficult task, and solutions in R and many other languages are easily Googleable - the point of this exercise is again to get used to the workflow and tools: RStudio, RMarkdown, Git/GitHub, Travis, etc.

2. Use your FizzBuzz function to display the correct output for a variety of inputs, make sure to show what happens for both good and bad input. 

**Solution**
```{r}
fizzbuzz <- function(x) {
  if (is.numeric(x) == FALSE) {
    print("not valid input, please try again")
    return (NULL)
  }
  for (i in 1:length(x)) {
    if (x[i] %% 1 != 0){
      print ("not valid input, please try again")
      return (NULL)
    }
    if (x[i] %% 15 == 0) {print("FizzBuzz")}
    else if (x[i] %% 3 == 0) {print("Fizz")}
    else if (x[i] %% 5 == 0) {print("Buzz")}
    else print(x[i])
  }
}
fizzbuzz(c(1, 3, 5, 15, 16))
fizzbuzz(1:15)
fizzbuzz("bad input")
fizzbuzz(c(13.45, 15))
fizzbuzz(c("bad input", 15))
```
My approach to the problem: A function fulfilling the requirement in this question need to be constructed and to react to different list of numbers of input. In order to be able to print a list of numbers, a (for loop) is used for R to be able to print out the list as desired. As of requirements for the list, the first requirement is that we would like the input to be only integer numbers, so that character & non-integer input will be invalid which is fulfilled by if statement. Then the next requirement is the fizzbuzz requirement, that is multiples of 3 print as fizz, multiples of 5 print as buzz, multiples of 15 print as fizzbuzz. This is also fulfilled by (if...else) statement. After we constructed the function, we test the function with several input that includes with integers, non-integers and characters and conclude that the function is constructed successfully.

## Q3. Rcpp

1. Read the online book [_Rcpp For Everyone_](https://teuder.github.io/rcpp4everyone_en/) by Masaki E. Tsuda.

2. Re-do FizzBuzz test using Rcpp. 

**Solution**
```{bash, warning = FALSE}
cat fizzbuzz.cpp
```

```{r,warning=FALSE}
library("Rcpp")
sourceCpp('fizzbuzz.cpp')
```

```{r}
Q3_fizzbuzz <- function(x) {
  if (is.numeric(x) == FALSE) {
    print ("not valid input, please try again")
    return (NULL)
  }
  else print (RCPP_fizzbuzz(x))
}
Q3_fizzbuzz(c(1, 3, 5, 15, 16))
Q3_fizzbuzz(1:15)
Q3_fizzbuzz("bad input")
Q3_fizzbuzz(c(13.45, 15))
Q3_fizzbuzz(c("bad input", 15))
```

